---
title: "Tuning the penalty"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
```

This case study chooses `pen` for `PELT` with the `"L2"` cost. It runs `PELT` over a grid of penalty values, records the number of change-points for each, and picks a `pen` where that number is stable.

This is useful when you would rather not use the methods in [Model selection](model-selection.html), or when the method has no elbow plot. `PELT` only returns the segmentation for a given `pen`, so it has no `$getHistory()` or `$plotElbow()`.

```{r setup, message = FALSE}
library(rupturesRcpp)
```

We simulate a series with three changes in mean, at 100, 200 and 300.

```{r}
set.seed(1)
tsMat = as.matrix(c(rnorm(100, 0), rnorm(100, 4), rnorm(100, 1), rnorm(100, 5)))

PELTObj = PELT$new(minSize = 1L, jump = 1L, costFunc = costFunc$new("L2"))
PELTObj$fit(tsMat)
```

## Number of change-points against `pen`

The grid is evenly spaced on a log scale. It starts at 5, which is small enough to let noise in. It ends at the cost of the whole series with no change-point, given by `$eval()`. With a larger `pen`, any change-point costs more than it can save, so `PELT` returns none.

```{r}
maxPen = PELTObj$eval(0, nrow(tsMat))
pens = exp(seq(log(5), log(maxPen), length.out = 50))
nCps = sapply(pens, function(pen) length(PELTObj$predict(pen = pen)) - 1)
```

`$predict()` returns the end-points, including `n`, so the number of change-points is one less than its length. The object is fitted once, so each `pen` only needs a new `$predict()`.

```{r, fig.width = 10, fig.height = 4, fig.alt = "Step plot of the number of change-points against pen on a log scale. The count falls quickly, stays at 3 over a wide range of pen, then drops to 1 and 0."}
plot(pens, nCps, type = "s", log = "x",
     xlab = "pen (log scale)", ylab = "Number of change-points")
```

A small `pen` lets noise through as extra change-points. A very large `pen` drops the true ones. In between, the number of change-points stays the same over a wide range of `pen`. That plateau is what we look for.

## Picking `pen` from the plateau

We take the longest run of equal counts, and the `pen` in the middle of it.

```{r}
runs = rle(nCps)
i = which.max(runs$lengths)
last = cumsum(runs$lengths)[i]
first = last - runs$lengths[i] + 1
penChosen = pens[(first + last) %/% 2]
c(nCps = runs$values[i], from = pens[first], to = pens[last], pen = penChosen)
```

The plateau is at 3 change-points, the true number. `PELT` is exact, so every `pen` on the plateau gives the same change-points.

```{r}
PELTObj$predict(pen = penChosen)
```

```{r, message = FALSE, fig.alt = "Simulated series split into four shaded segments at the change-points found by PELT with the chosen penalty."}
PELTObj$plot(main = paste0("method: PELT; costFunc: L2; pen: ", round(penChosen)))
```

The longest run is a simple rule. Look at the plot too: a short plateau, or two of similar length, means the data do not point to one clear number of change-points.
