---
title: "Change in mean and variance"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
```

This case study detects a change in mean and variance in a simulated two-dimensional series, with the `"SIGMA"` cost and binary segmentation (`binSeg`).

```{r setup, message = FALSE}
library(rupturesRcpp)
```

To demonstrate the package usage, we first consider a simple 2d time series with two piecewise Gaussian regimes and varying variance.

```{r}
set.seed(1)
tsMat = cbind(c(rnorm(100,0), rnorm(100,5,5)),
              c(rnorm(100,0), rnorm(100,5,5)))
```

As our example involves regimes with varying variance, a suitable `costFunc` option is `"SIGMA"`.  Since the segmentation objects' interfaces are similar, it is sufficient to demonstrate the usage of `binSeg` only.

```{r}
SIGMAObj = costFunc$new("SIGMA", addSmallDiag = TRUE, epsilon = 1e-6)
binSegObj = binSeg$new(minSize = 1L, jump = 1L, costFunc = SIGMAObj)
binSegObj$fit(tsMat)
```

Once fitted, `$predict()` and `$eval()` can be used. To view the configurations of the `binSeg` object, we can use `$describe()`.

```{r}
binSegObj$describe(printConfig = TRUE)
```

To obtain an estimated segmentation, we can use the `$predict()` method and specify a non-negative penalty value `pen`, which should be properly tuned. This returns a sorted integer vector of end-points, including the number of observations by design.

Here, we set `pen = 100` by hand. See [Tuning the penalty](case-study-pen.html) for how `pen` could be tuned.

```{r}
binSegObj$predict(pen = 100)
```

After running `$predict()`, the segmentation output is temporarily saved to the `binSeg` object, allowing users to use the `$plot()` method without specifying `endPts`.

```{r, message = FALSE, fig.height = 5, fig.alt = "Two simulated series split into two shaded segments at the detected change-point, t = 100."}
binSegObj$plot(d = 1:2,
               main = "method: binSeg; costFunc: SIGMA; pen: 100")
```

Instead of setting `pen`, we can choose the number of change-points with the elbow method (see [Model selection](model-selection.html#elbow-method)). `$plotElbow()` plots the cost against the number of change-points.

```{r, fig.height = 3.5, fig.alt = "Elbow plot of total cost against the number of change-points, dropping sharply at one change-point and then falling only slowly."}
binSegObj$plotElbow(maxK = 10)
```

The cost drops sharply at one change-point and then falls only slowly, so we take `nBkps = 1`. This gives the same change-point as `pen = 100`.

```{r}
binSegObj$predict(nBkps = 1)
```
