---
title: "Case studies"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
```

This page collects longer worked examples. The two below use simulated data and `binSeg`: a change in mean and variance (`"SIGMA"` cost), and a change in autoregressive dynamics (`"VAR"` cost). The second example also shows how to change an existing object through its active bindings.

```{r setup, message = FALSE}
library(rupturesRcpp)
```

## Simulated data

### 2-regime SIGMA example via binary segmentation

To demonstrate the package usage, we first consider a simple 2d time series with two piecewise Gaussian regimes and varying variance.

```{r}
set.seed(1)
tsMat = cbind(c(rnorm(100,0), rnorm(100,5,5)),
              c(rnorm(100,0), rnorm(100,5,5)))
```

As our example involves regimes with varying variance, a suitable `costFunc` option is `"SIGMA"`.  Since the segmentation objects' interfaces are similar, it is sufficient to demonstrate the usage of `binSeg` only.

```{r}
SIGMAObj = costFunc$new("SIGMA", addSmallDiag = TRUE, epsilon = 1e-6)
binSegObj = binSeg$new(minSize = 1L, jump = 1L, costFunc = SIGMAObj)
binSegObj$fit(tsMat)
```

Once fitted, `$predict()` and `$eval()` can be used. To view the configurations of the `binSeg` object, we can use `$describe()`.

```{r}
binSegObj$describe(printConfig = TRUE)
```

To obtain an estimated segmentation, we can use the `$predict()` method and specify a non-negative penalty value `pen`, which should be properly tuned. This returns a sorted integer vector of end-points, including the number of observations by design.

Here, we set `pen = 100`.

```{r}
binSegObj$predict(pen = 100)
```

After running `$predict()`, the segmentation output is temporarily saved to the `binSeg` object, allowing users to use the `$plot()` method without specifying `endPts`.

```{r, message = FALSE, fig.height = 5, fig.alt = "Two simulated series split into two shaded segments at the detected change-point, t = 100."}
binSegObj$plot(d = 1:2,
               main = "method: binSeg; costFunc: SIGMA; pen: 100")
```

### 2-regime VAR example

You can also modify a `binSeg` object's fields through its active bindings. To demonstrate this, we consider a piecewise vector autoregressive example with constant noise variance.

```{r}
set.seed(1)
tsMat = matrix(c(filter(rnorm(100), filter = 0.9, method = "recursive"),
                 filter(rnorm(100), filter = -0.9, method = "recursive")))
```

Here, the most suitable cost function is `"VAR"`. Instead of creating a new `binSeg` object, we will modify the current `binSegObj` as follows:

```{r}
VARObj = costFunc$new("VAR")
binSegObj$tsMat = tsMat
binSegObj$costFunc = VARObj
```

Modifying `tsMat` (or any other bindings) will automatically trigger `self$fit()` if a `tsMat` has already existed.

```{r}
binSegObj$describe(printConfig = TRUE)
```

We can then perform binary segmentation with `pen = 25` and plot the segmentation results.

```{r, message = FALSE, fig.alt = "Simulated autoregressive series with the change-points found by binSeg with the VAR cost marked by dashed lines."}
binSegObj$predict(pen = 25)
binSegObj$plot(d = 1L,
               main = "method: binSeg; costFunc: VAR; pen: 25")
```

The warning is expected: with `minSize = 1L`, some candidate segments have too few observations to fit the VAR model, so their cost falls back to an approximate solve.

## Real data

Case studies on real data are to be added.
