---
title: "Cost functions"
guide: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
```

This chapter lists the cost functions built into `rupturesRcpp`, their options, and how to supply your own cost as an R function. A cost function measures how badly a single model fits a segment; every segmentation class and `costFactory` takes one as a `costFunc` object.

```{r setup, message = FALSE}
library(rupturesRcpp)
```

## Built-in cost functions

Create a `costFunc` object by specifying the desired (supported) cost function and optional parameters:

```{r}
costFuncObj = costFunc$new("L2")
costFuncObj$pass() #output attributes corresponding to the specified cost function.
```

The following table shows the list of supported cost functions (pre-implemented ones are `PELT`-compatible; see [Segmentation methods](segmentation-methods.html)). Here, `n` is the segment length, and the time complexity is that of one segment query. For `"LinearL1"`, `p` is the number of features (columns of `tsMat`), `J` the number of regression coefficients (covariates plus the intercept) and `maxIter` the maximum number of IRLS iterations: every query refits the regression on the segment's rows, so it grows linearly in `n`.

| **Cost function** | **Description**                                                                                  | **Parameters/active bindings**           | **Dimension** | **Time complexity** |
|-------------------|--------------------------------------------------------------------------------------------------|------------------------------------------|----------------|----------------------|
| `"L1"`            | Sum of `L1` distances to the segment-wise median; robust to outliers.                            | `costFunc`                               | `multi`        | `O(nlog(n))`                 |
| `"L2"`            | Sum of squared `L2` distances to the segment-wise mean; faster but less robust than `L1`.        | `costFunc`                               | `multi`        | `O(1)`                 |
| `"SIGMA"`         | Log-determinant of the empirical covariance (divided by `n`, no bias correction); models varying mean&variance.                           | `costFunc`, `addSmallDiag`, `epsilon`    | `multi`        | `O(1)`                 |
| `"LinearL1"`      | Sum of `L1` residuals from a linear regression model, fit via Iteratively Reweighted Least Squares (IRLS); robust to outliers. | `costFunc`, `intercept`, `tol`, `maxIter` | `multi`        | `O(p·n·J²·maxIter)` |
| `"LinearL2"`      | Sum of squared residuals from a linear regression model with constant noise variance.            | `costFunc`, `intercept`                  | `multi`        | `O(1)`                 |
| `"LinearSIGMA"`   | Log-determinant of the residual covariance (divided by `n`, no degrees-of-freedom correction) from a linear regression model; models varying noise covariance around a regression mean. | `costFunc`, `intercept`, `addSmallDiag`, `epsilon` | `multi`        | `O(1)`                 |
| `"VAR"`           | Sum of squared residuals from a vector autoregressive model with constant noise variance.        | `costFunc`, `pVAR`                       | `multi`        | `O(1)`                 |
| `"Custom"`        | User-defined cost, supplied as a plain R function (see *User-defined cost functions* below).    | `costFunc`, `evalFun`, `paramFun`        | `multi`        | depends on `evalFun`  |

Every covariance in the package, in the `"SIGMA"`/`"LinearSIGMA"` costs and in the `cov` returned by `$get_params()`/`$segments()`, is the biased maximum-likelihood estimate: divided by the segment length `n`, with no Bessel (`n - 1`) or degrees-of-freedom (`n - q` for `q` regression coefficients) correction.

If active binding `costFunc` is modified by assigning to `costFuncObj$costFunc` and the required parameters are missing, the default parameters will be used. This does not apply to `"Custom"`: there is no sensible default `evalFun`, so `$pass()`/`$fit()` will error until one is set (see below).
```{r}
costFuncObj$costFunc = "VAR"
costFuncObj$pass()
```

## User-defined cost functions

If none of the built-in cost functions fit, `costFunc = "Custom"` lets you supply
your own as a plain R function, with no C++ required. It takes two active bindings:

- `evalFun` (required): called as `evalFun(segment, a, b)`, where `segment` is the
  raw matrix of rows `(a+1):b` of the fitted `tsMat`, and `a`/`b` are the same
  0-indexed `(a,b]` bounds `$eval(a, b)` uses. Must return a single numeric value.
- `paramFun` (optional, default `NULL`): same calling convention, used by internal
  parameter reporting; if omitted, `"Custom"` simply reports no parameters.

Passing `a`/`b` through, not just `segment`, is what makes this more than a
convenience wrapper: `evalFun` can use them to align `segment` against any other
externally-captured, position-indexed data (e.g. an exogenous series or weight
vector) that the package itself is never told about. Because each call crosses
back into R, it is substantially slower per call than the built-in costs, so prefer
one of those when it fits.

For example, `"L2"` re-implemented as a `"Custom"` cost:

```{r}
myL2eval = function(segment, a, b){
  segment = as.matrix(segment)
  cm = colMeans(segment)
  sum(sweep(segment, 2, cm, FUN = "-")^2)
}

customCF = costFunc$new("Custom", evalFun = myL2eval)
customCF$pass()
```

### Risk of data mismatch

The segmentation logic of existing modules only depends on being able to compute the cost for an arbitrary segment `(a,b]`; it does not depend on how the data are stored. Therefore, with a custom cost function, a mismatch can occur if the function relies on external data that are not part of the object passed to `$fit()`. The [Custom cost functions](case-study-custom.html#risk-of-data-mismatch) case study shows an example.

## Using a cost function

A `costFunc` object is passed to a segmentation class or to `costFactory`:

- [Segmentation methods](segmentation-methods.html#classes): `binSeg`, `Window`, `PELT` and `Dynp` search for change-points under the chosen cost.
- [Segment costs and parameters](segment-costs-and-parameters.html#costs-without-a-detection-algorithm): `costFactory` evaluates segments you choose, without a detection.
- Case studies: [Change in mean and variance](case-study-sigma.html) (`"SIGMA"`), [Change in autoregressive dynamics](case-study-var.html) (`"VAR"`) and [Custom cost functions](case-study-custom.html) (`"Custom"`).
