---
title: "Getting started"
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
```

This page installs `rupturesRcpp` and runs a first detection, from data to plot.

## What the package does

`rupturesRcpp` detects change-points in multivariate time series. A change-point is a time at which the behaviour of a series changes: its mean, its variance, its autocorrelation, or its relationship with other variables. Detection is offline: the whole series is available at once, and the goal is the set of change-points that best explains it.

## Installation {.tabset .tabset-pills}

Choose how to install. This site documents the development version.

### Development version (recommended)

r-universe builds the development version from the `main` branch on GitHub, with binaries for Windows, macOS and Linux, so no compiler is needed.

```{r, eval = FALSE}
install.packages("rupturesRcpp",
                 repos = c("https://edelweiss611428.r-universe.dev",
                           "https://cloud.r-project.org"))
```

### From source

Building from GitHub requires a C++ toolchain: Rtools on Windows; the Xcode Command Line Tools on macOS (`xcode-select --install`) and a GNU Fortran compiler (see the [CRAN macOS tools](https://cran.r-project.org/bin/macosx/tools/) page); or a C++ compiler on Linux (for example, `r-base-dev` on Debian and Ubuntu).

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("edelweiss611428/rupturesRcpp")
```

### CRAN release

```{r, eval = FALSE}
install.packages("rupturesRcpp")
```

The CRAN release (1.0.3) has `binSeg`, `Window` and `PELT` with the `"L1"`, `"L2"`, `"SIGMA"`, `"VAR"` and `"LinearL2"` costs, but not the rest of what this site uses:

- `Dynp` and `costFactory`;
- the `"LinearSIGMA"`, `"LinearL1"` and `"Custom"` costs;
- `$predict(nBkps = ...)`, `$getHistory()`, `$plotElbow()` and `$segments()`.

## A first detection

Load the package and check that the version is 2.0.0 or later:

```{r setup, message = FALSE}
library(rupturesRcpp)
packageVersion("rupturesRcpp")
```

Both series below change in mean and variance at $t = 100$. `tsMat` has one row per time point and one column per feature.

```{r}
set.seed(1)
tsMat = cbind(c(rnorm(100, 0), rnorm(100, 5, 5)),
              c(rnorm(100, 0), rnorm(100, 5, 5)))
```

The variance changes as well as the mean, so the `"SIGMA"` cost, which models both, suits this data. Binary segmentation (`binSeg`) is a fast search method.

```{r}
binSegObj = binSeg$new(costFunc = costFunc$new("SIGMA"))
binSegObj$fit(tsMat)
binSegObj$predict(pen = 100)
```

`$predict()` returns the end of each segment, so the last value is always the number of observations. Here it finds the one change-point at $t = 100$. The penalty `pen` sets how many change-points are kept.

```{r, message = FALSE, fig.height = 5, fig.alt = "Two simulated series split into two shaded segments at the detected change-point, t = 100."}
binSegObj$plot(d = 1:2)
```

## Next steps

The [Documentation](documentation.html) explains each part of a detection: the cost function, the segmentation method and how many change-points to keep. [Case studies](case-studies.html) has longer worked examples.
