---
title: "Model selection"
guide: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
```

This chapter shows how to choose the number of change-points without tuning `pen` directly: by reading the cost trajectory (the elbow method), or by asking `Dynp` for the exact optimum for each count.

```{r setup, message = FALSE}
library(rupturesRcpp)
```

## Elbow method

`binSeg` and `Window` both build up their segmentation by adding one change-point
at a time: `binSeg` by recursively splitting the segment that most reduces
cost, `Window` by ranking candidate local maxima by gain. `$getHistory()`
exposes that trajectory directly, so you can inspect it, or choose the number
of change-points via the "elbow method", instead of only tuning `pen`.

The examples below use the piecewise VAR series from [Change in autoregressive dynamics](case-study-var.html):

```{r}
set.seed(1)
tsMat = matrix(c(filter(rnorm(100), filter = 0.9, method = "recursive"),
                 filter(rnorm(100), filter = -0.9, method = "recursive")))

binSegObj = binSeg$new(minSize = 1L, jump = 1L, costFunc = costFunc$new("VAR"))
binSegObj$fit(tsMat)
head(binSegObj$getHistory())
```

The warning is expected: with `minSize = 1L`, some candidate segments have too few observations to fit the VAR model, so their cost falls back to an approximate solve.

`$plotElbow()` renders this as a `ggplot` object (cost vs. number of
change-points); look for where the marginal decrease in cost flattens out to
pick `k`.

```{r, fig.height = 3.5, fig.alt = "Elbow plot of total cost against the number of change-points, dropping sharply at one change-point and then levelling off."}
binSegObj$plotElbow(maxK = 10)
```

Once a `k` is chosen, `$predict()` accepts it directly via `nBkps`, which
takes precedence over `pen` when both are supplied:

```{r}
binSegObj$predict(nBkps = 1)
```

`nBkps` is treated as an upper bound, not a strict requirement: for `binSeg`
it returns its own best answer among the splits it already explored, and for
`Window` the `nBkps` highest-gain local maxima it found. Neither is
guaranteed to be the *globally* optimal segmentation for that count. `Window`
in particular can only ever offer as many change-points as it found local
maxima for; if you ask for more, `$predict()` returns what's available and
reports the shortfall via a message rather than erroring.

## Exact segmentation with Dynp

The caveat above (`binSeg`/`Window` returning their own best answer for a given `nBkps`, not necessarily the *globally* optimal one) is exactly what `Dynp` addresses. It finds the segmentation that is exactly optimal for a specified number of change-points, by building a full dynamic-programming table instead of searching greedily. The cost is complexity: the table takes `O(nBkpsMax * M^2)` time, where `M` is the number of `(minSize, jump)`-admissible positions, versus `binSeg`'s near-`O(n log n)` greedy search or `PELT`'s pruned search for a penalty.

```{r}
set.seed(1121)
signals = as.matrix(c(rnorm(100, 0, 1), rnorm(100, 5, 1)))

DynpObj = Dynp$new(minSize = 1L, jump = 1L) # nBkpsMax left NULL -> resolved at $fit()
DynpObj$fit(signals)
```

`$describe()` shows the two fields specific to `Dynp`: the user-set `nBkpsMax` (here left `NULL`) and the `resolvedNBkpsMax` actually used, capped at 20 by default.

```{r}
DynpObj$describe(printConfig = TRUE)
```

`$predict()`, `$eval()`, and `$plot()` all work exactly as with `binSeg`/`Window`/`PELT`, either via `nBkps` or via a penalty:

```{r}
DynpObj$predict(nBkps = 1)
```

```{r}
DynpObj$predict(pen = 100) # same split, reached via a penalty instead
```

```{r, message = FALSE, fig.alt = "Series split into two shaded segments at the change-point t = 100 found by Dynp."}
DynpObj$plot(main = "method: Dynp; costFunc: L2; nBkps: 1")
```

Since `$fit()` already computes the exact minimal cost for *every* change-point count from `0` to `nBkpsMax` in one pass (exposed via `$costPath()`), `$getHistory()`/`$plotElbow()` come for free here; no extra search is needed the way `binSeg`/`Window` would need to keep growing their own nested sequence:

```{r}
head(DynpObj$getHistory())
```

Unlike `binSeg`/`Window`, there is no `added_bkp` column: each `k`'s solution is independently exact and need not be nested inside the solution for `k+1`, so "the one breakpoint added at this step" is not generally well-defined. Use `$predict(nBkps = k)` to get the full breakpoint set for a given `k`.
